### 2019-03-20 papers/haskell19/relatedwork.tex
Added template with base structure and a bunch of code

\documentclass[conference, fleqn]{IEEEtran}

%% ODER: format ==         = "\mathrel{==}"
%% ODER: format /=         = "\neq "
%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stmaryrd}
\DeclareFontFamily{OT1}{cmtex}{}
\DeclareFontShape{OT1}{cmtex}{m}{n}
  {<5><6><7><8>cmtex8
   <9>cmtex9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
\DeclareFontShape{OT1}{cmtex}{m}{it}
  {<-> ssub * cmtt/m/it}{}
\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
  {<5><6><7><8>cmtt8
   <9>cmbtt9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
\DeclareFontShape{OT1}{cmtex}{bx}{n}
  {<-> ssub * cmtt/bx/n}{}
\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU

\newcommand{\Sp}{\hskip.33334em\relax}

\newcommand{\Conid}[1]{\mathit{#1}}
\newcommand{\Varid}[1]{\mathit{#1}}
\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
\newcommand{\plus}{\mathbin{+\!\!\!+}}
\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
\newcommand{\sequ}{\mathbin{>\!\!\!>}}
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}
\usepackage{polytable}

%mathindent has to be defined
\@ifundefined{mathindent}%
  {\newdimen\mathindent\mathindent\leftmargini}%
  {}%

\def\resethooks{%
  \global\let\SaveRestoreHook\empty
  \global\let\ColumnHook\empty}
\newcommand*{\savecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
\newcommand*{\restorecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
\newcommand*{\aligncolumn}[2]%
  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}

\resethooks

\newcommand{\onelinecommentchars}{\quad-{}- }
\newcommand{\commentbeginchars}{\enskip\{-}
\newcommand{\commentendchars}{-\}\enskip}

\newcommand{\visiblecomments}{%
  \let\onelinecomment=\onelinecommentchars
  \let\commentbegin=\commentbeginchars
  \let\commentend=\commentendchars}

\newcommand{\invisiblecomments}{%
  \let\onelinecomment=\empty
  \let\commentbegin=\empty
  \let\commentend=\empty}

\visiblecomments

\newlength{\blanklineskip}
\setlength{\blanklineskip}{0.66084ex}

\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
\let\hspre\empty
\let\hspost\empty
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

\EndFmtInput
\makeatother
%
%
%
%
%
%
% This package provides two environments suitable to take the place
% of hscode, called "plainhscode" and "arrayhscode". 
%
% The plain environment surrounds each code block by vertical space,
% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
% similar to formulas. Note that if these dimensions are changed,
% the spacing around displayed math formulas changes as well.
% All code is indented using \leftskip.
%
% Changed 19.08.2004 to reflect changes in colorcode. Should work with
% CodeGroup.sty.
%
\ReadOnlyOnce{polycode.fmt}%
\makeatletter

\newcommand{\hsnewpar}[1]%
  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}

% can be used, for instance, to redefine the code size, by setting the
% command to \small or something alike
\newcommand{\hscodestyle}{}

% The command \sethscode can be used to switch the code formatting
% behaviour by mapping the hscode environment in the subst directive
% to a new LaTeX environment.

\newcommand{\sethscode}[1]%
  {\expandafter\let\expandafter\hscode\csname #1\endcsname
   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}

% "compatibility" mode restores the non-polycode.fmt layout.

\newenvironment{compathscode}%
  {\par\noindent
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed\)%
   \par\noindent
   \ignorespacesafterend}

\newcommand{\compaths}{\sethscode{compathscode}}

% "plain" mode is the proposed default.
% It should now work with \centering.
% This required some changes. The old version
% is still available for reference as oldplainhscode.

\newenvironment{plainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newenvironment{oldplainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

% Here, we make plainhscode the default environment.

\newcommand{\plainhs}{\sethscode{plainhscode}}
\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
\plainhs

% The arrayhscode is like plain, but makes use of polytable's
% parray environment which disallows page breaks in code blocks.

\newenvironment{arrayhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\parray}%
  {\endparray\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newcommand{\arrayhs}{\sethscode{arrayhscode}}

% The mathhscode environment also makes use of polytable's parray 
% environment. It is supposed to be used only inside math mode 
% (I used it to typeset the type rules in my thesis).

\newenvironment{mathhscode}%
  {\parray}{\endparray}

\newcommand{\mathhs}{\sethscode{mathhscode}}

% texths is similar to mathhs, but works in text mode.

\newenvironment{texthscode}%
  {\(\parray}{\endparray\)}

\newcommand{\texths}{\sethscode{texthscode}}

% The framed environment places code in a framed box.

\def\codeframewidth{\arrayrulewidth}
\RequirePackage{calc}

\newenvironment{framedhscode}%
  {\parskip=\abovedisplayskip\par\noindent
   \hscodestyle
   \arrayrulewidth=\codeframewidth
   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
   \hline\framedhslinecorrect\\{-1.5ex}%
   \let\endoflinesave=\\
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \framedhslinecorrect\endoflinesave{.5ex}\hline
   \endtabular
   \parskip=\belowdisplayskip\par\noindent
   \ignorespacesafterend}

\newcommand{\framedhslinecorrect}[2]%
  {#1[#2]}

\newcommand{\framedhs}{\sethscode{framedhscode}}

% The inlinehscode environment is an experimental environment
% that can be used to typeset displayed code inline.

\newenvironment{inlinehscode}%
  {\(\def\column##1##2{}%
   \let\>\undefined\let\<\undefined\let\\\undefined
   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
   \def\fromto##1##2##3{##3}%
   \def\nextline{}}{\) }%

\newcommand{\inlinehs}{\sethscode{inlinehscode}}

% The joincode environment is a separate environment that
% can be used to surround and thereby connect multiple code
% blocks.

\newenvironment{joincode}%
  {\let\orighscode=\hscode
   \let\origendhscode=\endhscode
   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
   %\let\SaveRestoreHook=\empty
   %\let\ColumnHook=\empty
   %\let\resethooks=\empty
   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
  {\origendhscode
   \global\let\hscode=\orighscode
   \global\let\endhscode=\origendhscode}%

\makeatother
\EndFmtInput
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% HRep

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Constructor names

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% gen

% Local Variables:
% TeX-mbigstarer: "main.lhs.tex"
% TeX-command-default: "Make"
% End:

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{todonotes}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{mathdots}
\usepackage{yhmath}
\usepackage{cancel}
\usepackage{color}
\usepackage{siunitx}
\usepackage{array}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{bm}
\usepackage{microtype}
\usepackage{pgfplots}
\usetikzlibrary{fadings}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros
\newcommand{\tocite}{\textbf{CITE}}
\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
\newcommand{\dragenp}{\textbf{DRAGEN\!+}\xspace}

\newcommand{\hrep}[1]{\ensuremath{H\!Rep_{#1}}}
\newcommand{\gen}[1]{\ensuremath{gen_{#1}}}
\newcommand{\term}[1]{\ensuremath{#1^\bigstar}}
\newcommand{\evalrep}[2]{ \ensuremath{\llbracket #1 \rrbracket_{#2}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some parameters
\setlength{\mathindent}{\parindent}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title
\title{Random Generation of Rich \\ Abstract Data Type Values }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Authors
\author{
  \IEEEauthorblockN{
    % 1\textsuperscript{st}
    Agust\'in Mista
  }
  \IEEEauthorblockA{
    % \textit{Department of Computer Science and Engineering} \\
    \textit{Chalmers University of Technology}\\
    Gothenburg, Sweden \\
    mista@chalmers.se
  }
\and
\IEEEauthorblockN{
  % 2\textsuperscript{nd}
  Alejandro Russo
  }
  \IEEEauthorblockA{
    % \textit{Department of Computer Science and Engineering} \\
    \textit{Chalmers University of Technology}\\
    Gothenburg, Sweden \\
    russo@chalmers.se
  }
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstract

\begin{abstract}

  % Status
  Automatic generation of random values described by abstract data types (ADTs)
  is often a hard task.
  % data types (ADTs), % ---specially when considering ADTs which
  % represent HTML pages, images, programs or other structural rich objects.
  % Situation
  State-of-the-art random testing tools often can automatically synthesize
  random data generators based on ADTs definitions.
  %
  In that manner, generated values comply with the structure described by
  ADTs---something that proves useful when testing software which expects
  complex input like web pages, images, or even programs.
  %
  % Problem
  However, it sometimes becomes necessary to generate structural richer ADTs
  values in order to test deeper software layers.
  %
  % Solution
  In this light, we propose to leverage static information found in the
  codebase as a manner to improve the generation process.
  %
  Namely, our generators are capable to consider how programs branch on input
  data as well as how ADTs values are built via interfaces.
  %
  % Concretely, we propose to consider not only ADTs definitions, but also how
  % programs branch on input data as well as how ADTs values get manipulated via
  % interfaces.
  %
  We implement a tool, called {\dragenp}, responsible to synthezise generators
  for ADTs values while providing compile-time guarantees about their
  distributions.
  %
  Our compile-time predictions allow {\dragenp} to provide an heuristic that
  tries to adjust the distribution of generators to what developers might want.
  \todo[inline,author=Ale]{Case study}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Keywords
\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Sections

\section{Introduction}

% Scenario of random testing
Random testing is a promising approach to finding bugs
\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
%
\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
Haskell community.
%
It requires developers to specify (i) \emph{testing properties} which describe
programs behavior and (ii) \emph{random data generator} based on the
\underline{\emph{types}} of the expected inputs (e.g., an integer, an string,
etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
%
QuickCheck then generates random test cases and reports those violating testing
properties.
%

% Situation with Quick Check, ADTs, tools
QuickCheck comes equipped with random generators for built-in types, while it
requires to manually write generators for user-defined abstract data types
(ADTs).
%
Recently, there have been a proliferation of tools to automatically derive
QuickCheck generators for ADTs
\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
%
The main difference about these tools lies on the guarantees provided to ensure
\emph{the termination of the generation process} and the \emph{distribution of
  random values}.
%
Despite their differences, these tools guarantee that generated values are
\emph{well-typed}.
%
In other words, generated values follow the structure described by the
definition of the ADT.
%

%% How to use random generated ADTs
Well-typed ADT values are specially useful when testing programs which expect
highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
%
Generating ADT values also proves fruitful when looking for vulnerabilities with
fuzzers \cite{GriecoCB16,grieco2017}.
%
%% Problem
%
Despite these success stories, ADT type-definitions do not often capture all the
invariants expected from the data that they are intended to model.
%
As a result, even if random values are well-typed, they might not present enough
structure to penetrate into deep layers of software.

%% Our proposal
In this work, we propose a novel improvement in the generation process of ADT
values by exploiting some static information found in the codebase.
%
More specifically, to refine the structure of generated values, we propose a
generation process that is capable to consider how programs pattern-matched on
ADTs values as well as how they get manipulated via interfaces.
%
Furthermore, we show how to predict (at compile time) the distribution of the
\emph{expected} numbers of ADT constructors, values fitting a certain pattern,
and calls to interfaces.
%
For that, we extend some recent results on applying \emph{branching
  processes}---a simple stochastic model conceived to study population
growth---to predict the distribution of QuickCheck
generators\cite{DBLP:conf/haskell/MistaRH18}.
%
We implement our ideas in a tool, called {\dragenp}, that is capable to
automatically synthesize QuickCheck generators for ADT values, where the
distributions of random values can be adjusted at compile-time to what
developers might want based on our predictions.
%
\todo[inline,author=Ale]{Add here later about test cases when we know what they
  are}

We remark that, although this work focuses on Haskell algebraic data types, this
technique is general enough to be applied to most programming languages.
%with some level of support for composite types.

% The main contribuitions of this paper are:
% %
% \begin{itemize}
%   %
% \item We identify two patological scenarios for which standard type-driven
%   automatic derivation tools fail to synthesize practical random generators, due
%   to a lack of either type structure or domain knowleadge (Section
%   \ref{sec:randomtesting}).
%   %
% \item We present a generation technique able to encode stronger properties of
%   the target data type by reifying the static information present on the program
%   codebase (Section \ref{sec:hrep}).
%   %
% \item We apply and extend the theory of branching processes to analitically
%   predict the average distribution of generated values.
%   %
%   Furthermore, we use the predictions to perform simulation-based optimization
%   of the random generation parameters (Section \ref{sec:synthesis}).
%   %
% \item We provide an implementation of our ideas in the form of a Haskell library
%   to perform automatic derivation of random generators capable to extract useful
%   structure information from the user source code.
%   %
% \end{itemize}

% Local Variables:
% TeX-master: "main.lhs.tex"
% TeX-command-default: "Make"
% End:
%\newpage

\section{Random Data Generation in Haskell} \label{sec:randomtesting}

In this section we briefly introduce the common approach for automatically
deriving random data generators in Haskell using a type-driven approach, along
with its main drawbacks.

Haskell is a strongly typed programming language with a powerful type system.
%
It lets programmers encode a considerable amount of information about the
structure of their systems using data types that can be checked at compilation
time.
%
One of its key aspects is the support for Algebraic Data Types (ADTs).
%
Essentially, an ADTs is a composite type defined by combining other types in
terms of \textbf{sums} (also known as \emph{variant types}) and
\textbf{products} (or tuples) of other data types.
%
To exemplify this, consider the following type definition to encoding \ensuremath{\Conid{Html}}
expressions:
%
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{data}\;\Conid{Html}{}\<[12]%
\>[12]{}\mathrel{=}{}\<[16]%
\>[16]{}\Conid{Text}\;{}\<[22]%
\>[22]{}\Conid{String}{}\<[E]%
\\
\>[12]{}\,\ \vert\;{}\<[16]%
\>[16]{}\Conid{Sing}\;{}\<[22]%
\>[22]{}\Conid{String}{}\<[E]%
\\
\>[12]{}\,\ \vert\;{}\<[16]%
\>[16]{}\Conid{Join}\;{}\<[22]%
\>[22]{}\Conid{Html}\;{}\<[30]%
\>[30]{}\Conid{Html}{}\<[E]%
\\
\>[12]{}\,\ \vert\;{}\<[16]%
\>[16]{}\Conid{Tag}\;{}\<[22]%
\>[22]{}\Conid{String}\;{}\<[30]%
\>[30]{}\Conid{Html}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

In the previous definition, we declare \ensuremath{\Conid{Html}} as the sum of four possible
constructions: \ensuremath{\Conid{Text}} represents plain text values. \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Tag}} represent
singular and paired Html tags, and \ensuremath{\Conid{Join}} concats two expression one after
another.
%
We only encode a very small subset of the actual Html specification for
illustrative reasons.
%
In Haskell, \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}} are known as \emph{data
  constructors} (or constructors for short) and are used to distiguish which
variant of the data type we are constructing.
%
Each data constructor is defined as a product of zero or more types known as
\emph{fields}.
%
For instance, \ensuremath{\Conid{Text}} has a field of type \ensuremath{\Conid{String}}, whereas \ensuremath{\Conid{Join}} two recusive
fields of type \ensuremath{\Conid{Html}}.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this reprentation, the expression ``\ensuremath{<\!\!html\!\!>\!\!hello\!\!<\!\!hr\!\!>\!\!bye\!\!<\!\!/html\!\!>}'' can be encoded as:
%
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Conid{Tag}\;\text{\ttfamily \char34 html\char34}\;(\Conid{Join}\;(\Conid{Join}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}(\Conid{Text}\;\text{\ttfamily \char34 hello\char34})\;(\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}))\;(\Conid{Text}\;\text{\ttfamily \char34 bye\char34})){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
%
Additionally, we can define a function \ensuremath{\Varid{render}} to serialize \ensuremath{\Conid{Html}} values as
follows:
%
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{render}\mathbin{::}\Conid{Html}\to \Conid{String}{}\<[E]%
\\
\>[B]{}\Varid{render}\;(\Conid{Text}\;\Varid{t})\mathrel{=}\Varid{t}{}\<[E]%
\\
\>[B]{}\Varid{render}\;(\Conid{Sing}\;\Varid{t})\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
\\
\>[B]{}\Varid{render}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{render}\;\Varid{x}\plus \Varid{render}\;\Varid{y}{}\<[E]%
\\
\>[B]{}\Varid{render}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}\plus \Varid{render}\;\Varid{x}\plus \text{\ttfamily \char34 </\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

In the previous definition, \ensuremath{\Varid{render}} is described using \emph{pattern matching}
over each possible kind of value.
%
Using pattern matching we can define functions idiomatically by defining
different function clauses for each input pattern we are interested on.
%
Patterns can be defined to match specific constructors, literal values or
variable subexpressions (like \ensuremath{\Varid{t}}, \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} in the definition of \ensuremath{\Varid{render}}).
%
They can also be nested in order to match very specific patterns of values.

\subsection*{\textbf{Type-Driven Generation of Random Values}}

In order to generate random value of types involving user defined ADTs, most
approaches require the user to provide a random data generator for each one of
them.
%
This is a cumbersome and error prone task that closely follows the data type
structure.
%
For instance, consider the following definition of a \quickcheck random
generator for the type \ensuremath{\Conid{Html}}:
%
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
\column{6}{@{}>{\hspre}c<{\hspost}@{}}%
\column{6E}{@{}l@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}gen_{Html}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{}[\mskip1.5mu {}\<[6E]%
\>[9]{}(\mathrm{2},{}\<[14]%
\>[14]{}\Conid{Text}{}\<[20]%
\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{},{}\<[6E]%
\>[9]{}(\mathrm{1},{}\<[14]%
\>[14]{}\Conid{Sing}{}\<[20]%
\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary})\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\hsindent{4}{}\<[4]%
\>[4]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{}[\mskip1.5mu {}\<[6E]%
\>[9]{}(\mathrm{2},{}\<[14]%
\>[14]{}\Conid{Text}{}\<[20]%
\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{},{}\<[6E]%
\>[9]{}(\mathrm{1},{}\<[14]%
\>[14]{}\Conid{Sing}{}\<[20]%
\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{},{}\<[6E]%
\>[9]{}(\mathrm{3},{}\<[14]%
\>[14]{}\Conid{Join}{}\<[20]%
\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Html}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html}){}\<[E]%
\\
\>[4]{}\hsindent{2}{}\<[6]%
\>[6]{},{}\<[6E]%
\>[9]{}(\mathrm{4},{}\<[14]%
\>[14]{}\Conid{Tag}{}\<[20]%
\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html})\mskip1.5mu]){}\<[E]%
\ColumnHook
\end{hscode}\resethooks
 %$
%
This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
paremetrize the generation process up to an external natural number known as the
\emph{generation size}.
%
This parameter is chosen by the user, and we use to limit the maximum amount of
recursive calls that this random generator can perform.
%
When called with a positive generation size, this generator can pick to generate
among any \ensuremath{\Conid{Html}} data constructor of with a explicitly given frequency that can
be chosen by the user (2, 1, 3 and 4 for \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}}, \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}}
constructors, respectively).
%
When it picks to generate a \ensuremath{\Conid{Val}} or a \ensuremath{\Conid{Sing}} data constructor, it also
generates a random \ensuremath{\Conid{String}} value using the standard overloaded generator
\ensuremath{\Varid{arbitrary}} (\quickcheck provides standard random generators for most base types
like \ensuremath{\Conid{String}}, \ensuremath{\Conid{Int}}, \ensuremath{\Conid{Bool}}, etc.). \footnote{%
  The operators \ensuremath{\mathop{\mathsmaller{\langle \$ \rangle}}} and \ensuremath{\mathop{\mathsmaller{\langle \ast \rangle}}} are used in Haskell to combine values obtained
  from effectful computations (like calling to a random generator) and they are
  not particularly relevant for the point being made in this work.}
%$
On the other hand, when it picks to generate either an \ensuremath{\Conid{Join}} constructor, it
also generates two independent random subexpression recursively, decreasing the
generation size by a unit on each recursive invocation (\ensuremath{\Varid{smaller}\;gen_{Html}}).
%
The case of random generation of \ensuremath{\Conid{Tag}} constructors follows analogously.

This random process keeps calling itself recursively until the generation size
reaches zero, where the generator is constrained to pick only among terminal
data constructors, being \ensuremath{\Conid{Text}} and \ensuremath{\Conid{Sing}} the only possible choices in our
particular case.

% Strictly decreasing the the generation size by one on each recursive call
% results in a useful property over the generated data: every generated value has
% at most |n| levels, where |n| is the generation size set by the user.
%
% This property enables us to model the generation process using the theory of
% branching processes introduced by \tocite, and extended in this work as
% described in Section \ref{sec:synthesis}.

The previous definition is rather mechanical, except perhaps for the chosen
generation frequencies.
%
In this light, it easy to extend this procedure to any data type defined in an
algebraic fashion.
%
Fortunately, there exists different meta-programming tools to avoid the user
from having to mechanically write random generators over and over again for each
user-defined ADT.
%
The simplest tool for such purpose is \megadeth \tocite.
%
Given a target data type, it synthesizes a random generator for it that behaves
similarly to the one presented above, where the generation frequencies are
defined to be uniform across constructors.
%
However, picking among different data constructors with uniform frequency can
lead to a generation process biased towards generating (in average) very small
values, regardless of the generation sized set by the user \tocite.

\dragen is meta-programming tool conceived to mitigate this problem.
%
Instead of setting a uniform generation probability of data constructors, this
tool this tool uses the theory of branching processes to modelize and predict
analitically the average distribution of data constructors generated on each
random value.
%
This prediction mechanism is used to feedback a simulation-based optimization
process that adjusts the generation frequency of each data constructor in order
to obtain a particular distribution of values that can be specified by the user,
providing this way a more flexible testing environment while still being mostly
automated.

Althogh both \megadeth and \dragen synthesize random generators that are
theoretically capable to generate the whole space of values of the target data
type. the limitations arise quickly when we consider that the underlying
generation model is essentially the same: they pick a single data constructor
and recusively generate each required subexpression independently.
%
In practice, this procedure is often too generic to generate random data with
enough structural complexity to be used for testing purposes.

% In this work we identify two sources of additional structure information which
% are not considered by the aforementioned automatic derivation tools to obtain
% better random data generators:

In this work we identify two patological situations which are not properly
handled by the aforementioned derivation tools:

\begin{enumerate}
\item The target code behaves differently over inputs matching specific patterns
  of nested values.
\item The target code encodes a significant portion of its structure on its
  abstract interface.
\end{enumerate}

In Section \ref{sec:hrep} we show how this structural information can be used to
synthesize richer random generators automatically.
%
We proceed to exemplify the previous points in detail.

\subsection*{\textbf{Presence of Complex Pattern Matchings}}

To exemplify the first problematic scenario, suppose we want to use randomly
generated \ensuremath{\Conid{Htmls}}s to test a property comprising the following function:
%
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{simpl}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
\\
\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\mathrel{=}\Conid{Text}\;(\Varid{t1}\plus \Varid{t2}){}\<[E]%
\\
\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\mathrel{=}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y}))){}\<[E]%
\\
\>[B]{}\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Join}\;(\Varid{simpl}\;\Varid{x})\;(\Varid{simpl}\;\Varid{y}){}\<[E]%
\\
\>[B]{}\Varid{simpl}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x})\mathrel{=}\Conid{Tag}\;\Varid{t}\;(\Varid{simpl}\;\Varid{x}){}\<[E]%
\\
\>[B]{}\Varid{simpl}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

This function simplifies sequences of \ensuremath{\Conid{Text}} constructors into a single big
\ensuremath{\Conid{Text}} constructor.
%
To do so, it has to pattern match against sequences of \ensuremath{\Conid{Text}} constructors
combined by a \ensuremath{\Conid{Join}} constructor using nested patterns (see \ensuremath{\Varid{simpl}} fist and
second clauses).
%
The remaining clauses are only meant to propagate this simplification within
nested expressions.

>>>>

\section{Related Work}
\label{sec:relatedwork}
